Swarm and Evolutionary Based Algorithms used for Optimization

### 2023-01-21 release/code/latex/includes/header.tex
Release Added

\section{Results}
\label{sec:results}
\subsection{Pressure Vessel Design (Original)}
\label{subsec:pressure_vessel_problem_original}

In this section are presented the results for the original variable boundings of Pressure Vessel Design,
as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:pressure_vessel_problem_original}:

\begin{table}[H]
\centering
\caption{Best Fits for Pressure Vessel Design (Original)}
\label{best_fits:pressure_vessel_problem_original}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &        $x_3$ &        $x_4$ &         $f_x$ \\
\midrule
       EP & 0.00000000 & 0.00000000 &  40.32041055 & 200.00000000 &    0.08150806 \\
       ES & 0.00000000 & 0.00000000 &  40.32554414 & 200.00000000 &    0.08161190 \\
       GA & 0.00337818 & 0.00106498 &  46.44471109 & 129.50295637 &   17.06920868 \\
    BeesA & 0.07617198 & 0.07091908 & 104.75191725 &  44.23570263 & 2393.70106213 \\
      FFA & 0.00000000 & 0.00000000 &  40.39658670 & 200.00000000 &    0.08306113 \\
      PSO & 0.00000000 & 0.00000000 &  40.31961791 & 200.00000000 &    0.08149204 \\
\bottomrule
\end{tabular}}
\end{table}

While EP, ES, FFA and PSO seems to give similar results, it is noticeable that the cost function
gave some discrepant results for GA and BeesA.

Additional information is given by the Table~\ref{function_values:pressure_vessel_problem_original},
where are summarized statistics from all 100 results given by different starting points (populations).

\begin{table}[H]
\centering
\caption{Statistical Information about function values for Pressure Vessel Design (Original)}
\resizebox{\columnwidth}{!}{%
\label{function_values:pressure_vessel_problem_original}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &         Mean F &       Median F &          Max F &       StdDev F \\
\midrule
       EP &    0.08150806 &     0.15642850 &     0.09093798 &     1.32637552 &     0.16807714 \\
       ES &    0.08161190 &   718.92910990 &     1.00278859 &  9443.54269260 &  2060.07774212 \\
       GA &   17.06920868 &   525.00734162 &   361.53879602 &  3112.03503082 &   454.52386703 \\
    BeesA & 2393.70106213 & 17564.33024022 & 15751.91467117 & 53850.38271384 & 10514.60780411 \\
      FFA &    0.08306113 &     0.13296127 &     0.13106271 &     0.20084847 &     0.02396876 \\
      PSO &    0.08149204 &     0.24191732 &     0.08159173 &    10.00611523 &     1.00072349 \\
\bottomrule
\end{tabular}}
\end{table}

From statistical viewpoint, no algorithm tested for this paper have stable solutions.

The Figure~\ref{fig:pressure_vessel_design_original_boxplot} gives a visual representation
for Table~\ref{function_values:pressure_vessel_problem_original}:

\begin{figure}[H]
\centering
\caption{Boxplot for Pressure Vessel Design (Original)}
\label{fig:pressure_vessel_design_original_boxplot}
\includegraphics[scale=0.5]{images/pressure_vessel_problem_original_boxplot.png}
\end{figure}

Again, as it can be seen on the boxplot, there are no stability guaranteed for solution
in any algorithms.

Figure~\ref{fig:pressure_vessel_problem_original_convergence} shows the
function maximum, minimum and mean function of algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design (Original)}
\label{fig:pressure_vessel_problem_original_convergence}
\includegraphics[width=0.4 \textwidth]{images/pressure_vessel_problem_original_convergence.png}
\end{figure}

All algorithms tested have slow evolution for the original variable boundings, as it is shown in the convergence evolution.

\subsection{Pressure Vessel Design}
\label{subsec:pressure_vessel_problem}

In this section are presented the results for a tweaked variable boundings of Pressure Vessel Design,
as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:pressure_vessel_problem}:

\begin{table}[H]
\centering
\caption{Best Fits for Pressure Vessel Design}
\label{best_fits:pressure_vessel_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &       $x_3$ &        $x_4$ &         $f_x$ \\
\midrule
       EP & 0.75000000 & 0.35000000 & 40.68318345 & 195.00000000 & 5534.57345496 \\
       ES & 0.75000000 & 0.35000000 & 40.68322151 & 195.00000000 & 5534.57917966 \\
       GA & 0.75000816 & 0.35003661 & 40.68379662 & 195.00019653 & 5534.83662424 \\
    BeesA & 0.75000000 & 0.35000000 & 40.68319525 & 195.00000000 & 5534.57516657 \\
      FFA & 0.75000000 & 0.35000000 & 40.68318772 & 195.00000000 & 5534.57401617 \\
      PSO & 0.75000000 & 0.35000000 & 40.68318348 & 195.00000000 & 5534.57345480 \\
\bottomrule
\end{tabular}}
\end{table}

It is noticeable in this table that the best results and the best minimizers are quite the same,
so it cannot be said that all algorithms will solve the problem from this standpoint.

Another view of the solutions are presented in Table~\ref{function_values:pressure_vessel_problem}, where
are summarized the statistics of all 100 possible solutions for different starting points.

\begin{table}[H]
\centering
\caption{Statistical Information about function values for Pressure Vessel Design}
\resizebox{\columnwidth}{!}{%
\label{function_values:pressure_vessel_problem}
\begin{tabular}{lrrrrr}
\toprule
Algorithm &         Min F &        Mean F &      Median F &         Max F &   StdDev F \\
\midrule
       EP & 5534.57345496 & 5534.63426098 & 5534.61416291 & 5534.92993607 & 0.05804548 \\
       ES & 5534.57917966 & 5535.28165367 & 5535.06149254 & 5538.03003895 & 0.65098467 \\
       GA & 5534.83662424 & 5538.36788889 & 5536.92932748 & 5550.91639342 & 3.57079951 \\
    BeesA & 5534.57516657 & 5537.97495231 & 5537.04824684 & 5548.55587234 & 3.28239740 \\
      FFA & 5534.57401617 & 5534.58849395 & 5534.58435493 & 5534.63296149 & 0.01341777 \\
      PSO & 5534.57345480 & 5534.57381593 & 5534.57355270 & 5534.57625290 & 0.00053848 \\
\bottomrule
\end{tabular}}
\end{table}

It is hard to see any statistical  difference looking at any value in the table except standard deviation.
GA and BeesA seems to have the worst solutions, and EP, FFA and PSO seems to have a well defined behavior
since they have the lowest standard deviations, that is, all the solutions don't spread too much from each
other.

A quick way to access the Table~\ref{function_values:pressure_vessel_problem} is shown on Figure~\ref{fig:pressure_vessel_design_boxplot}.

\begin{figure}[H]
\centering
\caption{Boxplot for Pressure Vessel Design}
\label{fig:pressure_vessel_design_boxplot}
\includegraphics[scale=0.5]{images/pressure_vessel_problem_boxplot.png}
\end{figure}

It shows the same information as the Table~\ref{function_values:pressure_vessel_problem}
and it evidences that not all algorithms have the same mean, so it is a clear evidence
that not all algorithm give similar results from an arbitrary starting point (population).

Figure~\ref{fig:pressure_vessel_problem_convergence} shows the
function maximum, minimum and mean function of algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Pressure Vessel Design}
\label{fig:pressure_vessel_problem_convergence}
\includegraphics[width=0.4 \textwidth]{images/pressure_vessel_problem_convergence.png}
\end{figure}

EP, ES and PSO quickly converges all the solution once it find a viable minimum, whe GA, BeesA and FFA
maintain exploration and exploitation for all the epochs.
From Table~\ref{function_values:pressure_vessel_problem},
PSO has the best solution stability for the proud-and-lame-homemade custom bounding problem.

\subsection{Spring Tension Design}
\label{ssubec:spring_problem}

In this section are presented the results of Spring Tension Design optimization,
as exposed in Section~\ref{subsubsec:methodology-spring-tension-design}.

The best results for all algorithms are presented in Table~\ref{best_fits:spring_problem}:

\begin{table}[H]
\centering
\caption{Best Fits for Spring Tension Design}
\label{best_fits:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
Algorithm &      $x_1$ &      $x_2$ &       $x_3$ &      $f_x$ \\
\midrule
       EP & 0.05078716 & 0.34120356 & 11.93931414 & 0.01250534 \\
       ES & 0.05000000 & 0.32292194 & 13.26929635 & 0.01252956 \\
       GA & 0.05260626 & 0.38419911 &  9.63922108 & 0.01252305 \\
    BeesA & 0.05141008 & 0.35564113 & 11.06109433 & 0.01250200 \\
      FFA & 0.05148403 & 0.35770175 & 10.95787946 & 0.01249803 \\
      PSO & 0.05000000 & 0.32241990 & 13.31368880 & 0.01252626 \\
\bottomrule
\end{tabular}}
\end{table}

Although the best fit points have noticeable differences, the cost function $f_x$ have similar results, that is, the
function $f_x$ seems to have weak minimal points.

Another view of the solutions are presented in Table~\ref{function_values:spring_problem}, where
are summarized the statistics of all 100 possible solutions for different starting points (populations).

\begin{table}[H]
\centering
\caption{Statistical Information about function values for Spring Tension Design}
\label{function_values:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrrr}
\toprule
Algorithm &      Min F &     Mean F &   Median F &      Max F &   StdDev F \\
\midrule
       EP & 0.01250534 & 0.01255960 & 0.01253156 & 0.01453933 & 0.00020190 \\
       ES & 0.01252956 & 0.01281857 & 0.01274283 & 0.01346214 & 0.00024181 \\
       GA & 0.01252305 & 0.01628575 & 0.01616768 & 0.02561808 & 0.00273276 \\
    BeesA & 0.01250200 & 0.01268010 & 0.01261691 & 0.01406679 & 0.00021629 \\
      FFA & 0.01249803 & 0.01254780 & 0.01253499 & 0.01268116 & 0.00003152 \\
      PSO & 0.01252626 & 0.01390866 & 0.01318774 & 0.03036537 & 0.00342240 \\
\bottomrule
\end{tabular}}
\end{table}

Looking at standard deviation, EP, ES and BeesA appear to have the same distribution;
FFA has the lowest function value spread around the mean; GA and PSO have the largest
function values spread around the mean and particularly GA tend to give discrepant results
that the rest.

A quick way to access the Table~\ref{function_values:spring_problem} is
shown on Figure~\ref{fig:spring_tension_design_boxplot}.

\begin{figure}[H]
\centering
\caption{Boxplot for Spring Tension Design}
\label{fig:spring_tension_design_boxplot}
\includegraphics[scale=0.5]{images/spring_problem_boxplot.png}
\end{figure}

It shows the same information as the Table~\ref{function_values:spring_problem}
and it evidences that not all algorithms have the same mean, so it is a clear evidence
that not all algorithm give similar results from an arbitrary starting point (population).

Figure~\ref{fig:spring_problem_convergence} shows the
function maximum, minimum and mean function of algorithm's evolution:

\begin{figure}[H]
\centering
\caption{Convergence lines for Spring Tension Design}
\label{fig:spring_problem_convergence}
\includegraphics[width=0.4 \textwidth]{images/spring_problem_convergence.png}
\end{figure}

EP, ES and FFA have the best fitness.
All algorithms seems to converge the minimum line very quickly when
the algorithm find a minimal point.
Apart from EP and ES, the algorithms seems to maintain exploration and exploitation
until the time runs out.

\subsection{Significance Test}
\label{subsec:pressure_vessel_problem_significance_test}

The Friedman test is the non-parametric test used to compare related sample data, that is,
when the same individual is evaluated more than once.
The Friedman test does not use the data numbers directly, but the ranks occupied by
them after the sorting done for each group separately. After sorting, the hypothesis
of equality of the sum of the ranks of each group is tested.

Interpretation is as follows:
\begin{itemize}
    \item Assume two hypothesis about the data:
        \subitem $H_0$: The mean for each population is equal.
        \subitem $H_1$: At least one population mean is different from the rest.
    \item Given p-value, decide:
        \subitem p-value $\geq 0.05$: Reject $H_1$ with $95\%$ of confidence and accept null hypothesis;
        \subitem Accept $H_1$ otherwise.
\end{itemize}

The Table~\ref{friedman_test} resumes the Friedman's Chi-Squared Test made using
the 6 samples took by the algorithm's best results fore each run:

\begin{table}[H]
\centering
\caption{Significance Test Using Friedman Chi-Squared Test}
\label{friedman_test}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrr}
\toprule
                          Problem &   Rank &    p-value \\
\midrule
Pressure Vessel Design (Original) & 425.07428571 & 0.00000000 \\
           Pressure Vessel Design & 436.16000000 & 0.00000000 \\
            Spring Tension Design & 339.05714286 & 0.00000000 \\
\bottomrule
\end{tabular}}
\end{table}

Using the interpretation rules, since the p-value is less than $0.05$, the null hypotesis $H_0$
is rejected and each problem have at least one population with different mean from the rest.
Thus, not all algorithms gives the same result for the same problem.

Developed by F. Wilcoxon in 1945, the paired Wilcoxon test is based on the ranks
of intrapair differences. This non-parametric test, used to compare related samples,
is an alternative to the Student t-test when samples do not follow a normal distribution.
Therefore, the Wilcoxon test is used to test whether sample medians are equal in cases
where the assumption of normality is not satisfied or when it is not possible to check
this assumption.

Interpretation is as follows:
\begin{itemize}
    \item Assume two hypothesis about the data:
        \subitem $H_0$: the distributions of both samples are equal;
        \subitem $H_1$: the distributions of both samples are not equal.
    \item Given p-value, decide:
        \subitem p-value $\leq 0.05$: Reject $H_1$ with $95\%$ of confidence and accept null hypothesis;
        \subitem Accept $H_1$ otherwise.
\end{itemize}

The Table~\ref{wilcoxon_test:pressure_vessel_problem_original} gives the Wilcoxon test results
for \textit{Pressure Vessel Design} using the original variables bounding as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

\begin{table}[H]
\centering
\caption{Significance Test Using Wilcoxon Test for Pressure Vessel Design (Original)}
\label{wilcoxon_test:pressure_vessel_problem_original}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\toprule
  --- & EP & ES & GA & BeesA & FFA & PSO \\
\midrule
   EP & 0.0 & 1.20e-16 & 3.90e-18 & 3.90e-18 & 0.07 & 2.02e-4 \\
   ES & 1.20e-16 & 0.0 & 6.56-06 & 4.96e-18 & 5.49e-14 & 1.41e-17 \\
   GA & 3.90e-18 & 6.56e-06 & 0.0 & 3.90e-18 & 3.90e-18 & 3.90e-18 \\
BeesA & 3.90e-18 & 4.96e-18 & 3.90e-18 & 0.0 & 3.90e-18 & 3.90e-18 \\
  FFA & 0.07 & 5.49e-14 & 3.90e-18 & 3.90e-18 & 0.0 &  5.34e-05 \\
  PSO & 2.02e-4 & 1.41e-17 & 3.90e-18 &  3.90e-18 & 5.34e-05 & 0.0 \\
\bottomrule
\end{tabular}}
\end{table}

It is easy to notice that, from Wilcoxon test, FAA and EP have different distributions.
$H_1$ is only accepted in case of FFA and EP: they distributions seems to be not equal.

The Table~\ref{wilcoxon_test:pressure_vessel_problem} gives the Wilcoxon test results
for \textit{Pressure Vessel Design} using the proud-and-lame-homemade variables bounding
as exposed in Section~\ref{subsubsec:methodology-pressure-vessel-design}.

\begin{table}[H]
\centering
\caption{Significance Test Using Wilcoxon Test for Pressure Vessel Design}
\label{wilcoxon_test:pressure_vessel_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\toprule
  --- & EP & ES & GA & BeesA & FFA & PSO \\
\midrule
   EP & 0.0 & 9.04e-18 & 3.90e-18 & 6.12e-18 & 8.79e-12 & 4.40e-18 \\
   ES & 9.04e-18 & 0.0 & 1.71e-14 & 1.64e-13 & 4.81e-18 & 3.90e-18 \\
   GA & 3.90e-18 & 1.71e-14 & 0.0 & 0.38 & 3.90e-18 & 3.90e-18 \\
BeesA & 6.12e-18 & 1.64e-13 & 0.38 & 0.0 & 4.30e-18 & 3.90e-18 \\
  FFA & 8.79e-12 & 4.81e-18 & 3.90e-18 & 4.30e-18 & 0.0 & 4.02e-18 \\
  PSO & 4.40e-18 & 3.90e-18 & 3.90e-18 & 3.90e-18 & 4.02e-18 & 0.0 \\
\bottomrule
\end{tabular}}
\end{table}

EP, ES, FAA and PSO appears to have the same distributions.
However, BeesA and GA appears to have different distributions.
This fact confirms the results found in Table~\ref{function_values:pressure_vessel_problem}.

The Table~\ref{wilcoxon_test:spring_problem} gives the Wilcoxon test results
for \textit{Spring Tension Design} as exposed in Section~\ref{subsubsec:methodology-spring-tension-design}.

\begin{table}[H]
\centering
\caption{Significance Test Using Wilcoxon Test for Spring Tension Design}
\label{wilcoxon_test:spring_problem}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lllllll}
\toprule
  --- & EP & ES & GA & BeesA & FFA & PSO \\
\midrule
   EP & 0.0 & 1.47e-16 & 4.01e-18 & 4.37e-12 & 0.01 & 2.33e-16 \\
   ES & 1.47e-16 & 0.0 &  1.90e-17 & 2.12e-06 & 3.24e-17 & 7.98e-12 \\
   GA & 4.01e-18 &  1.90e-17 & 0.0 & 8.02e-18 & 4.27e-18 & 1.29e-11 \\
BeesA & 4.36e-12 & 2.12e-06 & 8.02e-18 & 0.0 & 3.01e-11 & 6.82e-15 \\
  FFA & 0.01 & 3.23e-17 & 4.27e-18 & 3.01e-11 & 0.0 & 9.32e-18 \\
  PSO & 2.33e-16 & 7.98e-12 & 1.29e-11 & 6.82e-15 & 9.32e-18 & 0.0 \\
\bottomrule
\end{tabular}}
\end{table}

From Wilcoxon Test, the are no difference between all algorithms since p-value
is less than $0.05$.

\section{Discussion}

In this paper were tested 3 evolutionary algorithms (EP, ES and GA) and 3
swarm based algorithms (BeesA, FFA and PSO), implemented by the \textit{MealPy}
package. The chosen test problem were \textit{Pressure Vessel Design} and
\textit{Spring Tension Design} because they are well studied engineering problems
and are founs in many, if not all, optimization books everywhere. Thus, there are
more results to compare.

Both problems were implemented in full form, that is, with constraints, using the scheme of
\textit{penalty functions}, as it is recomended by \textit{MealPy}'s documentation.

Surprisingly, the \textit{Pressure Vessel Design} problem was a tricky one to solve.
Many articles point out the boundings as described for original problem
in Section~\ref{subsubsec:methodology-pressure-vessel-design}.
But, for some unknown reason, the same variable bounding does not work for this
implementation, and it was necessary to tweak the search space to give at least
comparable results with the literature.
It is not intended to point out errors in modelling or blame anyone, it is only a mere
note to a misbehaviour in solving a specific problem.

The results gives a notion that there are no panacea or universal solution for problems that
nothing is known except the cost function and constraints.
One algorithm's success to solve a specific problem is not a guarantee to solve anything
and so on.

At first, the \textit{Spring Tension Design} is easier to solve and is well behaved in the
given variable bounsings. But, at it is noticed in the Table~\ref{best_fits:spring_problem},
the cost function $f_x$ seems to have a plateau of weak minimizers, that is, values that gives
near te same results. The statistical parameters of the cost function values at the best fits
, as well the Wilcoxon test results fot this problem shows, there are no difference between the
distribution of each algorithm solution, but the Friedman Test (and clearly the boxplot in
figure~\ref{fig:spring_tension_design_boxplot}) indicates that they have significant
differences in the mean value. So, there are some algorithms not fitted for the problem's solution.
The best fit at all can be obtained using FFA algorithm, according to this paper.

The \textit{Pressure Vessel Design} was solver using two bounding schemes to show discrepancies
between the solutions.
First, it was solved using the literature boundings: the behaviour of the
solutions seems bizarre and make it evident that or the literature is wrong or the implementation
used in this paper is wrong.
Again, it is not intended heve to blame anyone.
The second solution, using the proud-and-lame-homemade
version of variable boundings seems to behave well.
The same behaviour of \textit{Spring Tension Design} is noticed in the tests, both statistical
, Wincoxon and Friedman Tests.

From similarity tests, the tests gave the following interpretations:

\begin{itemize}
    \item The best strategies to solve \textit{Pressure Vessel Design} in its proud-and-lame-homemade version
    are EP, ES, FFA and PSO, due to its stability and results similarity. The best of all is FFA due to its
    lower standard deviation.
    \item No strategy tested in this paper were able to solve \textit{Pressurew Vessel Design} in its original
    version and the available tools chosen for implementation.
    \item The best strategies to solve \textit{Spring Tension Design} are EP, ES, BeesA, FFA and PSO. The best of
    all is FFA due to its lower standard deviation.
\end{itemize}

\end{document}

>>>>

\section{Results}
\label{sec:results}
